Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-06-12 LaTex/paper.tex
adapted formulas for distance measures

\textcolor{red}{Need for more explanations}

\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse quantile profile plots (each collection comparing to 7 expected quantile values), each with only one bin for the sake of visualization, for separate sets of exemplary probability density estimation and observed data combinations. Each shown line represents the percentage of probability density estimation and observed data combinations for which the observed data point should be above and below the quantile of the predicted PDF indicated by that line (for example, the 0.50 line, corresponding to the median, indicates that 50 percent of all probability density estimation and observed data combinations in a given set of data should fall above the line, and 50 percent should fall below the line), respectively. The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of probability density estimations. For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles.

\textcolor{red}{need to explain how one can see this}

In the example of a parametric negative binomial PDF estimation of section \ref{sec:pdfEstimation}, this would indicate that both the mean and the variance estimations as well as the choice of a negative binomial distribution as underlying PDF were correct. Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias and the center and center-right columns of shaded circles illustrate a case for which the mean estimation has a bias.

@@

The methods described so far allow a detailed qualitative evaluation of PDF predictions. However, in order to also quantify the quality of the PDF predictions, a measure of the deviation of the PDF predictions from the optimal outcome given the observed target data is needed. To achieve this, we compare the CDF histogram of the predicted PDFs with the uniform distribution and define a prediction accuracy interval between 0 and 1.

Several different methods can be used to compute the deviance between two probability distributions. Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a distance function defined between two probability distributions on a given metric space (also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, a measure of how one probability distribution diverges from a second expected probability distribution. The first has the advantage that, unlike the second, it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.

\textcolor{red}{Need to add some formulas.}

>>>>

\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse quantile profile plots (each collection comparing to 7 expected quantile values), each with only one bin for the sake of visualization, for separate sets of exemplary probability density estimation and observed data combinations. Each shown line represents the percentage of probability density estimation and observed data combinations for which the observed data point should be above and below the quantile of the predicted PDF indicated by that line (for example, the 0.50 line, corresponding to the median, indicates that 50 percent of all probability density estimation and observed data combinations in a given set of data should fall above the line, and 50 percent should fall below the line), respectively. The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of probability density estimations. For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles. Here, all shaded circles exactly fall on the line, indicating that both mean and variance estimations, as well as the choice of a negative binomial distribution as underlying PDF, were correct. Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias, and the center and center-right columns of shaded circles illustrate cases for which the mean estimation has a bias.

@@

The methods described so far allow a detailed qualitative evaluation of PDF predictions. However, in order to also quantify the quality of the PDF predictions, a measure of the deviation of the PDF predictions from the optimal outcome given the observed target data is needed. To achieve this, we compare the CDF histogram of the predicted PDFs with the uniform distribution and define a prediction accuracy interval between 0 and 1. Several different methods can be used to compute the deviance between two probability distributions. Good choices are the \textbf{1st Wasserstein distance} \cite{olkin1982}, the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, and the \textbf{Jensen-Shannon divergence}  \cite{dagan1997}.

The 1st Wasserstein distance, also known as earth mover'€™s distance (EMD), represents a distance function defined between two probability distributions on a given metric space, and for our purposes here can defined by

\begin{equation}
\text{EMD}(P, Q) = 2 \cdot \frac{\sum_{k=1}^N |F_P(x_k) - F_Q(x_k)|}{N},
\end{equation}

where $F_P(X)$ and $F_Q(X)$ are the CDFs of the two PDFs $P(X)$ and $Q(X)$, respectively, and $x_k$ denotes the average value of $X$ in bin $k$, with $X$ being divided in $N$ bins.

The Kullback-Leibler divergence is a measure of how one probability distribution diverges from a second expected probability distribution. For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-“Leibler divergence from Q to P is defined as

\begin{equation}
D_{\text{KL}}(P \parallel Q) = \sum _{k=1}^N P(x_k) \log \left({\frac{P(x_k)}{Q(x_k)}}\right),
\end{equation}

where $\log$ can be either base-2 or natural logarithm.

The Jensen-Shannon divergence (JSD) can be seen as a symmetrized and smoothed version of the Kullback-Leibler divergence:

\begin{equation}
D_{\text{JSD}}(P \parallel Q) = 0.5  \cdot (D_{\text{KL}}(P \parallel M) + D_{\text{KL}}(P \parallel M)),
\end{equation}

where $M = 0.5  \cdot (P + Q)$.

The EMD has the advantage over the mentioned divergences that it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
